---
title: "Commuting in LA"
subtitle: "An analysis of my time behind the wheel"
output: html_document
---

## Summary

Anyone living in LA will know that spending time behind the wheel is part of life in this part of the world. In this document I visualize some of the data I have collected over the last few months on my commuting times in the drive between Marina Del Rey and my office in Santa Monica.

## Setup

### Load Packages

First we will load in the packages we need including the excellent `googlesheets` as well as our own `redbull` package for chart formatting.

```{r load packages, message = FALSE}
library(googlesheets)
library(tidyverse)
library(magrittr)
library(redbull)
```

```{r authorise googlesheets, echo = FALSE, message = FALSE, eval=FALSE}
options(httr_oob_default = TRUE)
gs_auth()
gs_ls()
```

```{r}
source("R/get_commuting_data.R")
```

Connect to our Google Sheet and extract the data we need.

```{r extract data, message = FALSE}
# Commuting sheet key
commuting_sheet_key <- "1jCm2hkRckKBRtXgIXCEiGAjfN51rV6soV16TF7qjsxo"

comm_data <- get_commuting_data(commuting_sheet_key)

comm_data_clean <- clean_commuting_data(comm_data)

```

Ok so let's look at monthly commute time averages.

```{r commute time over time, message = F}
theme_set(theme_rb())

comm_data.tidy %>%
  ggplot(aes(x = Trip_Time, y = Trip_Duration_Minutes, colour = Direction, Direction)) +
  geom_point() +
  facet_wrap(~ Direction, scales = "free_x") +
  geom_smooth(se = F)


```

```{r, eval=F}
comm_data.tidy %>%
  ggplot(aes(x = Trip_Weekday, y = Trip_Duration_Minutes, fill = Direction)) + 
  geom_boxplot() + 
  facet_grid(~ Direction)

comm_data.tidy %>%
  ggplot(aes(x = Trip_Date, y = Trip_Duration_Minutes, colour = Direction)) +
  geom_point() + 
  geom_smooth(se = F)

comm_data.tidy %>%
  ggplot(aes(x = as.factor(lubridate::month(Trip_Date)), y = Trip_Duration_Minutes, fill = Direction)) + 
  geom_boxplot() + 
  facet_grid(~ Direction)


comm_data.tidy %>% mutate(m = lubridate::month(Trip_Date)) %>% View()


comm_data.mod <- comm_data.tidy %>% 
  select(Trip_Duration_Minutes, Trip_Weekday, Direction, Trip_Date, Trip_Time) %>%
  mutate(Trip_Month = as.factor(lubridate::month(Trip_Date))) %>%
  select(-Trip_Date, -Trip_Time)

commuting_mod <- lm(Trip_Duration_Minutes ~ ., data = comm_data.mod)

```

